---
title: "p8105_hw2_SC4934"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r chunk_library, results = "hide", message = FALSE}
library(tidyverse)
library(readxl)
```



# Problem 1

## import data

```{r}
trans_ent = 
  read_csv(
    "data/chs03.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```


```{r, chunk_skim, results = "hide", message = FALSE}
skimr::skim(trans_ent)
```

## description: 
the dataset `trans_ent` contains 1868 rows and 19 columns. There are 15 character, 2 logical, and 2 numeric variables. 
it includes variables: line, station_name, station_latitude, station_longitude, all the vairables that start with 
"route" (which are routes served), entry, exit_only, vending, entrance_type, and ada. 
I used `janitor::clean_names` to start the tidying process, which converts column names to lower snake case. I also used `col_types` to convert some columns to character format. I used `select` to select the variables I wanted to include, and used `mutate` to convert entry variable to logical type. 

this data is not tidy!


```{r }
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

After selecting only `station_name` and `line` and using the `distinct` function to obtain all unique combinations, we can see that there are 465 distinct stations.


```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

the chunk above filters the ADA compliant stations among all distinct stations. 
There are 84 ADA compliant stations. 






