---
title: "p8105_hw2_SC4934"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r chunk_library, results = "hide", message = FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```



# Problem 1

## import data

```{r}
trans_ent = 
  read_csv(
    "data/chs03.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```


```{r, chunk_skim, results = "hide", message = FALSE}
skimr::skim(trans_ent)
```

## description: 
the dataset `trans_ent` contains 1868 rows and 19 columns. There are 15 character, 2 logical, and 2 numeric variables. 
it includes variables: line, station_name, station_latitude, station_longitude, all the vairables that start with 
"route" (which are routes served), entry, exit_only, vending, entrance_type, and ada. 
I used `janitor::clean_names` to start the tidying process, which converts column names to lower snake case. I also used `col_types` to convert some columns to character format. I used `select` to select the variables I wanted to include, and used `mutate` to convert entry variable to logical type. 

this data is not tidy!

## distinct stations:

```{r, chunk_distint_station}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

After selecting only `station_name` and `line` and using the `distinct` function to obtain all unique combinations, we can see that there are 465 distinct stations.


## ADA compliance

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

the chunk above filters the ADA compliant stations among all distinct stations. 
There are 84 ADA compliant stations. 

## allow entrance

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

37.70% of station entrances/exits without vending allow entrance. 


## reformat data 

to reformat data so that route number and route name are distinct variables

```{r}
trans_ent <- trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name")
```


to determine how many distinct stations serve the A train, and of those how many are ADA compliant: 

```{r}
trans_ent %>% 
      filter(route_name == "A")  %>% 
      select(station_name, line) %>% 
      distinct
  
trans_ent %>% 
  filter(route_name == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct

```

60 distinct stations serve the A train, and of those 17 are ADA compliant. 


# Problem 2

## Import and clean Excel file

```{r}
trash_wheel = read_excel('data/Trash Wheel Collection Data.xlsx',
                                 range = 'A2:N549',
                                 sheet = "Mr. Trash Wheel",
                                 ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    type = "Mr Trash Wheel",
    sports_balls = as.integer(sports_balls),
    dumpster = as.character(dumpster),
  ) 


professor_wheel = read_excel('data/Trash Wheel Collection Data.xlsx',
                                 range = 'A2:M96',
                                 sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    type = "Professor Trash Wheel",
    dumpster = as.character(dumpster),
    year = as.character(year)
  ) 

```



## merge data
```{r, chunk_merge}
mr_proff_merge <- bind_rows(trash_wheel, professor_wheel) %>% 
  janitor::clean_names()
```

## description: 

there are 641 observations in the dataset. key variables include `dumpster`, which distinguishes dumpsters, and `type`, which shows whether the data belongs to mr. trash wheel or professor trash wheel. 

total weight of trash collected by professor trash wheel is `r sum(mr_proff_merge[which(mr_proff_merge$type == "Professor Trash Wheel"), "weight_tons"])` tons, while the total number of sports balls collected by mr. trash wheel in 2020 is 856


```{r, chunk_sports_balls}

mr_proff_merge %>% 
  filter(type == "Mr Trash Wheel", year == "2020") %>% 
  pull(sports_balls) %>% 
  sum

```


# Problem 3

## Import and clean

```{r}
pols_month = 
  read_csv(
    "data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>% 
  mutate(
    month = month.abb[month],
    month = str_to_lower(month),
    prez_dem = na_if(prez_dem, 0) ,
    prez_gop = na_if(prez_gop, 0)
  ) %>% 
  relocate(prez_dem, .after = prez_gop) %>% 
    pivot_longer(
    prez_gop:prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_drop_na = TRUE
  ) %>% 
  select(-value, -day)
  


snp = 
  read_csv(
    "data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) %>% 
  mutate(
    month = month.abb[month],
    year = ifelse(year > 49, year + 1900, year + 2000),
    month = str_to_lower(month)
  ) %>% 
  select(year, month, close)

unemployment = 
  read_csv(
    "data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployement"
  ) 

```


## merge

```{r}
pols_snp_merge =
  left_join(snp, pols_month) %>% 
  janitor::clean_names()
  
  
  
pols_snp_unemployment = 
  left_join(unemployment, pols_snp_merge)%>% 
  janitor::clean_names()
 


```


## descriptions:

`pols_month` : contains 822 obserations and 9 variables. describes the number of politicians who are democratic or republican at a given date/time. it now also contains `president`, which shows whether the 
president at the time was democratic or republican. 

`snp` : contains 787 observations of 3 variables (after cleaning, originally was 2 variables). it contains the year and month, and the closing values of the S&P stock index under the variable `close`.

`unemployement` : after cleaning, contains 816 observations of 3 variables. it contains year and month, and the percent unemployment under the variable `unemployment`.

`pols_snp_unemployment` : the final merged dataset contains `pols_month`, `snp`, and `unemployment`. it was joined on month and year, and shows the combined results from all 3 datasets. contains 1602 rows and 11 columns. the range of years is `r range(pols_snp_unemployment$year)`. key variables include `month`, `year`, `president`, `unemployment`, and `close`







